Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-12-26 manuscript.Rmd

Following a wave of democratization efforts across the Middle East and North
Africa, modest Syrian protests began in March 2011. The regime responded harshly
with imprisonments, detentions, censorship, and military operations as the
resistance to Assad's regime spread. The fighting escalated dramatically in
March 2012, after increased unrest in the western city of Homs. International
actors---including the Red Cross/Red Crescent Society, the United Nations
General Assembly, the U.N. Security Council, and individual world
leaders---worked toward cease-fires, negotiations, and concessions without much
success. While many high-level military and diplomatic officials defected,
numerous loyal forces continued the regime's repression, and the Syrian people
turned to social media to coordinate their activism and communicate to the
world. Between March and June 2012, millions of tweets were sent within
Syria, revealing the details of the conflict suppressed by the government. A full account of the history of the Arab Spring is outside of the scope of the current article; more information can be found in Brownlee (2013) [fix bib] and Lynch (2016) [fix bib].

@@

## Social Media and Social Movements
Massive online social cohesion can manifest in real-world social behavior.
Social media platforms enable online social cohesion by offering distributed
information-gathering and real-time information dissemination. Given its
accessibility and proficiency as an information exchange tool, Twitter has
provided social and political activists an opportunity to complement (Greijdanus et al., 2020) [fix bib] and not hinder participation in (Harlow & Harp, 2012; Milošević-Đorđević & Žeželj, 2017) [fix bib] real-world social mobilization. During the 2012 Italian protests for global economic change, tweeting was determined to be more effective locally in discussing real-world, real-time events than traditional mainstream media platforms [@vicari2013public]. As people increasingly rely on social media for local news and information, we must understand exactly how these micro-blogs connect to and reflect real-world action.

Along with Facebook, Twitter has been credited as essential to the Arab
Spring---both for protesters and for the government's response to them
[@comninos2011twitter]. The Arab Spring has been deeply examined within the
political science literature [e.g., @lutterbeck2013arab;
@koehler2017political], but to our knowledge, the direct *coupling* between
Syrian Twitter cohesion and international action has not been investigated  (though hashtag usage has been linked to next-day protests; Steinert-Threlkeld et al., 2015) [fix bib]. In fact, some have been skeptical about the effect of social media on collective action at all [@seay2014slacktivism].

@@

At the same time, the insulation of online social platforms allows researchers
to detail real-world social dynamics at unprecedented resolution and scale
[@paxton2017finding]. The digital traces left by individuals and collectively as
groups are regarded as a valuable source of information that can offer powerful
insights into states of social disequilibrium, uprisings, and escalations into
civil wars. By examining these conversations, researchers can study the degree
of cohesion between speakers and the alignment between that cohesion and their
collective political behavior, just as the degree of cohesion among real-world protestors in past movements has been captured in their shared sign messaging (Windsor, 2020) [fix bib].

>>>>

Following a wave of democratization efforts across the Middle East and North
Africa, modest Syrian protests began in March 2011. The regime responded harshly
with imprisonments, detentions, censorship, and military operations as the
resistance to Assad's regime spread. The fighting escalated dramatically in
March 2012, after increased unrest in the western city of Homs. International
actors---including the Red Cross/Red Crescent Society, the United Nations
General Assembly, the U.N. Security Council, and individual world
leaders---worked toward cease-fires, negotiations, and concessions without much
success. While many high-level military and diplomatic officials defected,
numerous loyal forces continued the regime's repression, and the Syrian people
turned to social media to coordinate their activism and communicate to the
world. Between March and June 2012, millions of tweets were sent within
Syria, revealing the details of the conflict suppressed by the government. A full account of the history of the Arab Spring is outside of the scope of the current article; more information can be found in [@brownlee2013syria] and [@lynch2016ne].

@@

## Social Media and Social Movements
Massive online social cohesion can manifest in real-world social behavior.
Social media platforms enable online social cohesion by offering distributed
information-gathering and real-time information dissemination. Given its
accessibility and proficiency as an information exchange tool, Twitter has
provided social and political activists an opportunity to complement [@greijdanus2020psychology] and not hinder participation in [@harlow2012collective; @milovsevic2017civic] real-world social mobilization. During the 2012 Italian protests for global economic change, tweeting was determined to be more effective locally in discussing real-world, real-time events than traditional mainstream media platforms [@vicari2013public]. As people increasingly rely on social media for local news and information, we must understand exactly how these micro-blogs connect to and reflect real-world action.

Along with Facebook, Twitter has been credited as essential to the Arab
Spring---both for protesters and for the government's response to them
[@comninos2011twitter]. The Arab Spring has been deeply examined within the
political science literature [e.g., @lutterbeck2013arab;
@koehler2017political], but to our knowledge, the direct *coupling* between
Syrian Twitter cohesion and international action has not been investigated  (though hashtag usage has been linked to next-day protests; @steinert2015online). In fact, some have been skeptical about the effect of social media on collective action at all [@seay2014slacktivism].

@@

At the same time, the insulation of online social platforms allows researchers
to detail real-world social dynamics at unprecedented resolution and scale
[@paxton2017finding]. The digital traces left by individuals and collectively as
groups are regarded as a valuable source of information that can offer powerful
insights into states of social disequilibrium, uprisings, and escalations into
civil wars. By examining these conversations, researchers can study the degree
of cohesion between speakers and the alignment between that cohesion and their
collective political behavior, just as the degree of cohesion among real-world protestors in past movements has been captured in their shared sign messaging [@Windsor2020].
